Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-11-04 paper/paper.tex
edit conclusion Tested everything on bennet.

We have simulated flattened cell shapes similar to those observed
experimentally by Mannik \emph{et al.}, with a stochastic and a
deterministic version of Huang's widely used reaction-diffusion model.
We find that the stochastic model predicts experimentally accurate
qualitative behavoir in both the wild-type, pill-shaped cells (as
found in previous studies~\cite{fange2006noise, varma2008min,
  kruse2007experimentalist}), as well as in the new, flattened,
irregularly shaped cells.  These flattened cells invariably show
spatially irregular formation of maxima that is also seen
experimentally.

We find that the deterministic model~\cite{huang2003dynamic}
predicts behavior that contradicts what is seen experimentally, namely
robust and periodic bipolar oscillation in irregularly shaped cells,
with a minor intermediate maximum appearing in shape \emph{B}.  Thus
we conclude that this method is inconsistent with experiment, and
should not be relied upon for predictions of the behavior of the MinD
system.

In order to understand the irregular MinD oscillations observed in
experiments with flattened cells, we simulated flattened ``stadium''
shapes in order to distinguish between the effect of flattening and
the effect of irregular shape.  We found only a moderate difference in
the coherence of bipolar oscillations within these stadium shapes and
the irregular shapes observed by Mannik \emph{et al.}, and have
observed visually the same qualitative spatially irregular formation
of maxima in both.  Thus we conclude that the disruption of the MinD
oscillations seen by Mannik \emph{et al.}  is largely due to the
flattening of the cell shape, as opposed to its irregularity.

>>>>

We find agreement between our stochastic simulations and experiments
showing disrupted bipolar MinD oscillation in asymmetric flattened
E. coli cells~\cite{mannik2012robustness}.  As observed
experimentally, our simulations predict that MinD maxima will form in
a spatially irregular sequence of locations.  This result builds upon
existing results showing that this model is effective in a variety of
wild-type and mutated cell shapes~\cite{fange2006noise, varma2008min,
  kruse2007experimentalist}, and reinforces that the stochastic
variation of Huang's 2003 model~\cite{fange2006noise,
  kerr2006division} has considerable predictive power beyond the
standard wild-type cell.

%
%% We have simulated flattened cell shapes similar to those observed
%% experimentally by Mannik \emph{et al.}, with a stochastic and a
%% deterministic version of Huang's widely used reaction-diffusion model.
%% %
%% We find that the stochastic model predicts experimentally accurate
%% qualitative behavoir in both the wild-type, pill-shaped cells (as
%% found in previous studies~\cite{fange2006noise, varma2008min,
%%   kruse2007experimentalist}), as well as in the new, flattened,
%% irregularly shaped cells.  These flattened cells invariably show
%% spatially irregular formation of maxima as has been observed
%% experimentally.

In contrast to the stochastic model, the original deterministic
version of the model~\cite{huang2003dynamic} predicts behavior that
contradicts experimental observations, namely robust and periodic
bipolar oscillation in irregularly shaped cells, with a minor
intermediate maximum appearing in shape \emph{B}.  Thus we conclude
that this method is inconsistent with experiment, and cannot be
relied upon for predictions of the behavior of the MinD system.
\fixme{Confirm others? What failures have already been seen? Is this
  the most dramatic?}

We find that flattened but regular and symmetric cells exhibit MinD
oscillations that are qualitatively similar to the oscillations
observed experimentally (and in our simulation) in irregular and
asymmetric cells.  This demonstrates that asymmetry is not required in
order to induce spatially irregular MinD oscillations.  In addition,
the temporal regularity of oscillations is only moderately affected by
irregularity in the shape of flattened cells.  Thus we conclude
that it is the flattening of the cells rather than their irregular
shape that primarily causes the spatial irregularity of maxima which
is observed by Mannik \emph{et al.}~\cite{mannik2012robustness}.
%% %
%% In order to understand the irregular MinD oscillations observed in
%% experiments with flattened cells, we simulated flattened ``stadium''
%% shapes in order to distinguish between the effect of flattening and
%% the effect of irregular shape.
%% %
%% We found only a moderate difference in the coherence of bipolar
%% oscillations within these stadium shapes and the irregular shapes
%% observed by Mannik \emph{et al.}, and have observed visually the same
%% qualitative spatially irregular formation of maxima in both.  Thus we
%% conclude that the disruption of the MinD oscillations seen by Mannik
%% \emph{et al.}  is largely due to the flattening of the cell shape, as
%% opposed to its irregularity.
